---
title: "Homework 4"
author: "clbenning"
date: "2025-02-05"
output: html_document
---

## Question 1

```{r}
x <- 1.1
a <- 2.2 
b <- 3.3
```

### Question 1a
```{r}
z <- x^{a^b}
print(z)
```

### Question 1b
```{r}
z <- {x^a}^b

print(z)
```

### Question 1c
```{r}
z <- 3*{x^3} + 2*{x^2} + 1

print(z)
```

## Question 2 

### Question 2a 

vec_a <- c(seq(1:8), seq(7,1))

print(vec_a)

[1] 1 2 3 4 5 6 7 8 7 6 5 4 3 2 1

### Question 2b

vec_b <- c(seq(1:5))

vec_b <- rep(x = vec_b, times = vec_b)

print(vec_b)

[1] 1 2 2 3 3 3 4 4 4 4 5 5 5 5 5

### Question 2c 

vec_c1 <- c(seq(1:5))

vec_c2 <- c(seq(5,1))

vec_c <- rep(x = vec_c2, times = vec_c1)

print(vec_c)

[1] 5 4 4 3 3 3 2 2 2 2 1 1 1 1 1


## Question 3 

coord <- runif(2)

print(coord)

[1] 0.6778340 0.3549398

polar_coord <- c(sqrt(coord[1]^2+coord[2]^2),atan(coord[2]/coord[1]))

print(polar_coord)

[1] 0.7651413 0.4823789

## Question 4
```{r}
queue <- c("sheep", "fox", "owl", "ant")
```
### Question 4a

```{r}
# The serpent arrives and gets in line
queue[5] <- "serpent"
print(queue)
```

### Question 4b

```{r}
# The sheep enters the Ark 
queue <- queue[-1]
print(queue)
```

### Question 4c

```{r}
# The donkey arrives and talks his way to the front of the line
queue <- c("donkey", queue)
print(queue)
```

### Question 4d

```{r}
# The serpent gets impatient and leaves 
queue <- queue[-5]
print(queue)
```

### Question 4e

```{r}
# The owl gets bored and leaves 
queue <- queue[-3]
print(queue)
```

### Question 4f

```{r}
# The aphid arrives and the ant invites him to cut in line.
queue <- c(queue[1:2], "aphid", queue[3])
print(queue)
```

### Question 4g

```{r}
# Finally, determine the position of the aphid in the line.
which(queue == "aphid")
```

## Question 5

```{r}
vec <- 1:100
vec <- which((vec%%2 != 0) & (vec%%3 != 0) & (vec%%7 != 0))
print(vec)
```






