---
title: "Homework_08"
author: "clbenning"
date: "2025-03-24"
output: html_document
---

# Question 1: Working with fake data ---- 
```{r}
library(ggplot2)
library(MASS)
```

### Read in a data vector 

# ```{r}
# z <- rnorm(n=3000,mean=0.2)
# z <- data.frame(1:3000,z)
# names(z) <- list("ID","myVar")
# z <- z[z$myVar>0,]
# str(z)
# summary(z$myVar)
# ```
# #### Plot a histogram of fake data 
# 
# ```{r}
# p1 <- ggplot(data=z, aes(x=myVar, y=..density..)) +
#   geom_histogram(color="black",fill="darkgoldenrod2",size=0.2) 
# print(p1)
# ````
# 
# #### Add empiracal density curve 
# ```{r}
# p1 <-  p1 +  geom_density(linetype="dotted",size=0.75)
# print(p1)
# ```
# 
# #### Get the maximum liklihood parameters for normal distribution 
# 
# ```{r}
# normPars <- fitdistr(z$myVar,"normal")
# print(normPars)
# str(normPars)
# normPars$estimate["mean"] 
# ```
# 
# #### Plot the normal probability density on the histogram 
# 
# ```{r}
# meanML <- normPars$estimate["mean"]
# sdML <- normPars$estimate["sd"]
# 
# xval <- seq(0,max(z$myVar),len=length(z$myVar))
# 
#  stat <- stat_function(aes(x = xval, y = ..y..), fun = dnorm, colour="red", n = length(z$myVar), args = list(mean = meanML, sd = sdML))
#  p1 + stat
# ```
# 
# #### Add the exponential, uniform, gamma, and beta probability density curves 
# 
# ```{r}
# expoPars <- fitdistr(z$myVar,"exponential")
# rateML <- expoPars$estimate["rate"]
# 
# stat2 <- stat_function(aes(x = xval, y = ..y..), fun = dexp, colour="blue", n = length(z$myVar), args = list(rate=rateML))
#  p1 + stat + stat2
# ```
# 
# ```{r} 
# stat3 <- stat_function(aes(x = xval, y = ..y..), fun = dunif, colour="darkgreen", n = length(z$myVar), args = list(min=min(z$myVar), max=max(z$myVar)))
# p1 + stat + stat2 + stat3
# ``` 
#  
# ```{r} 
# gammaPars <- fitdistr(z$myVar,"gamma")
# shapeML <- gammaPars$estimate["shape"]
# rateML <- gammaPars$estimate["rate"]
# 
# stat4 <- stat_function(aes(x = xval, y = ..y..), fun = dgamma, colour="brown", n = length(z$myVar), args = list(shape=shapeML, rate=rateML))
#  p1 + stat + stat2 + stat3 + stat4
# ``` 
# 
# ```{r} 
# pSpecial <- ggplot(data=z, aes(x=myVar/(max(myVar + 0.1)), y=..density..)) +
#   geom_histogram(color="grey60",fill="cornsilk",size=0.2) + 
#   xlim(c(0,1)) +
#   geom_density(size=0.75,linetype="dotted")
# 
# betaPars <- fitdistr(x=z$myVar/max(z$myVar + 0.1),start=list(shape1=1,shape2=2),"beta")
# shape1ML <- betaPars$estimate["shape1"]
# shape2ML <- betaPars$estimate["shape2"]
# 
# statSpecial <- stat_function(aes(x = xval, y = ..y..), fun = dbeta, colour="orchid", n = length(z$myVar), args = list(shape1=shape1ML,shape2=shape2ML))
# pSpecial + statSpecial
# ```


# Question 2: Rye Data ----

## This data is a compilation of performance metrics, agronomic practices, and soil properties across the eastern half of the United States. The dataset includes a total of 5,695 cereal rye biomass observations across 208 site-years between 2001–2022 and encompasses a wide range of agronomic, soil, and climate conditions. 

#### Read in data set

```{r}
library(readr)
ryedat <- read_csv("rye_data.csv")
View(ryedat)

ryedat$study_id <- as.factor(ryedat$study_id)
ryedat$site_id <- as.factor(ryedat$site_id)
```

# Plot histogram of rye data
```{r}
plot <- ggplot(data = ryedat, aes(x =  shoot_biomass_mean_kg_ha, y = ..density..)) +
  geom_histogram(color="black",fill="aquamarine3",size=0.2) +
  theme_classic()
plot
```

# Emipirical Density Curce 

```{r}
plot <- plot + geom_density(linetype="dotted",size = 2)
plot
```

# Maximum Liklihood Parameters for Normal 

```{r}
normPars <- fitdistr(ryedat$shoot_biomass_mean_kg_ha, "normal")
print(normPars)
str(normPars)
normPars$estimate["mean"]
```

#Normal Distribution 

```{r}
meanML <- normPars$estimate["mean"]
sdML <- normPars$estimate["sd"]

xval <- seq(0,max(ryedat$shoot_biomass_mean_kg_ha),len=length(ryedat$shoot_biomass_mean_kg_ha))

stat <- stat_function(aes(x = xval, y = ..y..), fun = dnorm, colour="red", n = length(ryedat$shoot_biomass_mean_kg_ha), args = list(mean = meanML, sd = sdML))

plot + stat
```

# Exponential Probability Density
```{r}
expoPars <- fitdistr(ryedat$shoot_biomass_mean_kg_ha,"exponential")
rateML <- expoPars$estimate["rate"]

stat2 <- stat_function(aes(x = xval, y = ..y..), fun = dexp, colour="blue", n = length(ryedat$shoot_biomass_mean_kg_ha), args = list(rate=rateML))

plot + stat + stat2
```

# Uniform Probability Density 

```{r}
stat3 <- stat_function(aes(x = xval, y = ..y..), fun = dunif, colour="darkgreen", n = length(ryedat$shoot_biomass_mean_kg_ha), args = list(min=min(ryedat$shoot_biomass_mean_kg_ha), max=max(ryedat$shoot_biomass_mean_kg_ha)))

plot + stat + stat2 + stat3
```

# Gamma Probability Density 

```{r}
gammaPars <- fitdistr(ryedat$shoot_biomass_mean_kg_ha,"gamma")
shapeML <- gammaPars$estimate["shape"]
rateML <- gammaPars$estimate["rate"]

stat4 <- stat_function(aes(x = xval, y = ..y..), fun = dgamma, colour="brown", n = length(ryedat$shoot_biomass_mean_kg_ha), args = list(shape=shapeML, rate=rateML))

plot + stat + stat2 + stat3 + stat4
```

