---
title: "Homework_06"
author: "clbenning"
date: "2025-02-19"
output: html_document
---

# Creating Fake Data Sets To Explore Hypotheses

#### In an experiment looking at the effects of applying fertilizer to a corn field, the null hypothesis would be: Nitrogen treated plots will have the same corn yields as plots that did not receive fertilizer. The alternative hypothesis would be: Plots receiving nitrogen fertilizer will have a different yield than unfertilized plots. 

#### Mean corn yields will be 20 tons/acre with a standard deviation of 2.

## Generate random data set

```{r}
dat <- rnorm(100, mean = 20, sd = 2)

dat <- data.frame(dat)

dat$Fert <- sample(c("Fertilized", "Unfertilized"))

colnames(dat) <- c("Yield", "Fert")

dat
```
## Run an ANOVA and test the hypothesis 

#### p-value is 0.059 which shows that there not quite enough evidence to reject the null hypothesis. 

```{r}
mod <- lm(Yield ~ Fert, data = dat)
anova(mod)
```
```{r}
library(ggplot2)
library(Rmisc)
library(multcompView)
dat_sum <- summarySE(dat, measurevar = "Yield", groupvars = "Fert", na.rm = TRUE)
  
dat_plot <- ggplot(dat_sum, aes(y = Yield, x = Fert, fill = Fert)) +
  geom_bar(position = "dodge", stat="identity") +
  geom_errorbar(aes(ymin = Yield-se, ymax = Yield+se), width = .2, position = position_dodge(.9)) +
  scale_fill_manual(values=c("cornflowerblue", "darkgoldenrod1"))+
  scale_y_continuous(limits=c(0,25)) +
  theme_bw()+
  theme(text = element_text(size = 20)) + 
  theme(legend.position = "right", strip.background = element_rect(colour="black", fill="white")) +
  theme(axis.text.x = element_text(vjust = 0.5, hjust=.5, size = 15)) + 
  labs(x="Treatment", y="Yield tons " ~ac^-1) +
  labs(title="Corn Yields") +
  theme(plot.title = element_text(hjust = 0.5))

dat_plot
```

## For Loops to adjust sample size 

#### There is no sample size change between 10 and 200 that creates a significant effect.
```{r}
results <- data.frame(sample_size = integer(),
                       treatment = character(),
                       p_value = numeric())

treatments = c("Fertilized", "Unfertilized")

for (i in seq(10, 200, by = 5)) {  # Increase sample size by 5

  temp_dat <- dat[sample(1:nrow(dat), i, replace = TRUE), ]

  mod <- lm(Yield ~ Fert, data = temp_dat)
  
    p_value <- summary(mod)$coefficients[2] 

    results <- rbind(results, data.frame(sample_size = i, treatment = treatments, 
                                         p_value = p_value))
  }
```


