---
title: "Homework_10"
author: "clbenning"
date: "2025-04-09"
output: html_document
---

```{r setup, include=FALSE}
library(tidytuesdayR)
library(tidyverse)
library(ggplot2)

dat <- tidytuesdayR::tt_load('2023-07-11')
dat <- tidytuesdayR::tt_load(2023, week = 28)

global_temps <- dat$global_temps

seasonal_global_temps <- global_temps %>%
  pivot_longer(cols = 16:19, 
               names_to = "Season")
seasonal_global_temps$Temp <- seasonal_global_temps$value

monthly_global_temps <- global_temps %>%
  pivot_longer(cols = 2:13, 
               names_to = "Month")
monthly_global_temps$Temp <- monthly_global_temps$value

nh_temps <- dat$nh_temps
sh_temps <- dat$sh_temps

hem_temps <- rbind(nh_temps, sh_temps)
hem_temps$Hemisphere <- rep(c("nh", "sh"), each = 144)

hem_temps <- hem_temps %>%
  pivot_longer(cols = 16:19, 
               names_to = "Season")

hem_temps$Temp <- hem_temps$value
```

# Ridge Plot

## Plot the Land-Ocean Temperature Index for the Northern (nh) and Southern (sh) Hemispheres by meteorogical seasons

```{r}
library(ggridges)

ridge <- ggplot(data = hem_temps, aes(x = Temp, y = Season, fill = Season)) +
  ggridges::geom_density_ridges() +
  ggridges::theme_ridges() +
  facet_wrap(~Hemisphere, scale = "fixed") +
  labs(title="Land-Ocean Temperature Index") +
  theme(plot.title = element_text(hjust = 0.5)) +
  labs(x = "Temperature (°C)", y = "Seasons")

ridge
```

# Bar Plot ----

## Plot the changes in Land-Ocean Temperature Index in both hemispheres across meteorogical seasons and decades 

```{r}
library(Rmisc)

# Summarise the data by decades 
hem_temps_decades <- hem_temps %>%
  mutate(Decade = floor(Year / 10) * 10)

hem_temp_sum <- summarySE(hem_temps_decades, measurevar = "Temp", groupvars = c("Decade", "Hemisphere", "Season"), na.rm = TRUE)

# Plot using a bar graph

bar <- ggplot(data = hem_temp_sum) +
  aes(x = Decade, y = Temp, fill = Hemisphere) + 
  geom_bar(stat = "identity") +
  scale_fill_manual(values=c("darkolivegreen4", "goldenrod2")) +
  facet_wrap(~Hemisphere, scale = "fixed") +
  labs(x="Decade", y="Temperature (°C)") +
  labs(title="Global Land-Ocean Temperature Index") +
  theme(plot.title = element_text(hjust = 0.5)) +
  theme_bw()

bar
```

## Plot the changes in Global Land-Ocean Temperature Index across the seasons and include bars for standard deviation 
```{r}
# Summarise the data by decades 
global_temps_decades <- seasonal_global_temps %>%
  mutate(Decade = floor(Year / 10) * 10)

global_temp_sum <- summarySE(global_temps_decades, measurevar = "Temp", groupvars = c("Decade", "Season"), na.rm = TRUE)

# Plot using a bar graph with standard deviation 

bar2 <- ggplot(data = global_temp_sum) +
  aes(x = Decade, y = Temp, fill = Season) + 
  geom_bar(stat = "identity") +
  geom_errorbar(aes(ymin = (Temp-sd), ymax = (Temp+sd), width=.2)) +
  scale_fill_manual(values=c("darkslategray3","goldenrod2","darkolivegreen4", "orangered3")) +
  facet_wrap(~Season, scale = "fixed") +
  labs(x="Decade", y="Temperature (°C)") +
  labs(title="Global Land-Ocean Temperature Index") +
  theme(plot.title = element_text(hjust = 0.5)) +
  theme_bw()

bar2
```

# Scatter Plot ----

## Plot the changes in temperature in both hemispheres and seasons with a linear regression line


scale_fill_manual(values=c("darkslategray3","goldenrod2","darkolivegreen4", "orangered3")) +

```{r}
Scatter <- ggplot(data = hem_temp_sum) +
  aes(x = Decade, y = Temp, color = Hemisphere) +
  geom_point(size = 3, stroke = 0.5) +
  geom_smooth(method="lm") +
  labs(x="Decade", y="Temperature (°C)") +
  labs(title="Global Land-Ocean Temperature Index") +
  theme(plot.title = element_text(hjust = 0.5)) +
  facet_wrap(~Season, scale = "fixed") +
  theme_bw()

Scatter
```
## Monthly scatter plots of changes in global temperature across the decades 
```{r}
# Summarise the data by decades 
monthly_global_temps_decades <- monthly_global_temps %>%
  mutate(Decade = floor(Year / 10) * 10)

month_order <- c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec")

monthly_global_temps_decades$Month <- factor(monthly_global_temps_decades$Month, 
                                              levels = month_order)

Scatter2 <- ggplot(data = monthly_global_temps_decades) + 
            aes(x = Decade,y = Temp) + 
            geom_point()+
  facet_grid(.~Month) +
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 60, vjust = 0.5, hjust=.5, size = 7))+
  labs(title="Monthly Global Land-Ocean Temperature Index") +
  theme(plot.title = element_text(hjust = 0.5)) +
  labs(x="Decade", y="Temperature (°C)")

Scatter2 
```

